Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-04 WSSSPE2_report/WSSSPE2_review.tex
Section 8: small changes.

This group, with just three papers but a large amount of interest and
participation from attendees, focused on the institutional, social, and cultural
mechanisms that encourage the creation and maintenance of shared software, in the
context of what now exists but focusing on what mechanisms are desired and how
we might achieve them.

@@

\subsubsection{First breakout discussion: Group A}
The first sub-group discussed issues around the current system for credit and
current incentives, which it called ``hacking the incentive structure.'' The
group considered four potential points of leverage:

@@

Third that we could change academic culture, rather than worrying about the
systems. This was mostly focused on citations, because they matter for hiring,
promotion, and tenure decisions. The group discussed how we could weigh the
citations within papers better than we now do? How could we identify the five
citations that really matter for a paper, distinguishing them from the related
works and general background that are also cited. Perhaps we could break these
out in the reference list, working with publishers to implement this. Or maybe
we could also break out categories of citations, such as the most important
software used, the previous publication that we are building from, the data that
we actually used, etc. The Moore Foundation's award in data science was given as
an example, asking proposers: What are the five canonical citations that are
most important to your work?~\cite{moore-canonical} This would be a way of
giving credit and assigning importance to these works, differently from how we
just count citations today. A possible action that the group discussed was
conducting a longitudinal study of most useful software, data, etc.\ in a
discipline.

@@

This group started by discussing who should be incentivized, thinking of two
categories of people: those in science (who could be incentivized to do better,
more shareable, more sustainable work), and those in industry but interested in
science (who could be incentivized to contribute to science.) It was pointed out
that we are not yet clear enough on exactly what we want to incentivize,
suggesting that we need to have a clearer picture of ``good computational work''
and what sort of contributions are truly generative for science.

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right'' way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors weren't necessarily following those suggestions
anyway)~\cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational'' and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing'' PIs to include their ``software
people'' on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

@@

In the remaining discussion sessions, the group chose to split into three
subgroups to discuss a version of these topics: {\em citation ecosystems}, {\em
taxonomy of contributorship/guidelines for software for tenure review}, and {\em
prizes}. The subgroups were asked to clearly identify
\begin{itemize}
\item the problem to be solved
\item steps towards a solution.
\end{itemize}

>>>>

This group, with just three papers but a high level of interest and
participation from attendees, focused on the institutional, social, and cultural
mechanisms that encourage the creation and maintenance of shared software, in the
context of what now exists, what mechanisms are desired, and how
we might achieve them.

@@

\subsubsection{First breakout discussion: Group A}
The first sub-group discussed issues around the current system for credit and
incentives, which it called ``hacking the incentive structure.'' The
group considered four potential points of leverage:

@@

Third that we could change academic culture, rather than worrying about the
systems. This was mostly focused on citations, because they matter for hiring,
promotion, and tenure decisions. The group discussed how we could weigh the
citations within papers better than we now do. How could we identify the five
citations that really matter for a paper, distinguishing them from the related
works and general background that are also cited. Perhaps we could break these
out in the reference list, working with publishers to implement this. Or maybe
we could also break out categories of citations, such as the most important
software used, the previous publication that we are building from, the data that
we actually used, etc. The Moore Foundation's award in data science was given as
an example, asking proposers: What are the five canonical citations that are
most important to your work?~\cite{moore-canonical} This would be a way of
giving credit and assigning importance to these works, differently from how we
just count citations today. A possible action that the group discussed was
conducting a longitudinal study of most useful software, data, etc.\ in a
discipline.

@@

This group started by discussing who should be incentivized, thinking of two
categories of people: those in science (who could be incentivized to do better,
more shareable, more sustainable work), and those in industry but interested in
science (who could be incentivized to contribute to science). It was pointed out
that we are not yet clear enough on exactly what we want to incentivize,
suggesting that we need to have a clearer picture of ``good computational work''
and what sort of contributions are truly generative for science.

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless, the discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right'' way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors were not necessarily following those suggestions
anyway)~\cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational'' and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing'' PIs to include their ``software
people'' on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

@@

In the remaining discussion sessions, the group chose to split into three
subgroups to discuss these topics: {\em citation ecosystems}, {\em
taxonomy of contributorship/guidelines for software for tenure review}, and {\em
prizes}. The subgroups were asked to clearly identify
\begin{itemize}
\item the problem to be solved
\item steps towards a solution.
\end{itemize}
